---
title: "Make posterior"
output: html_notebook
---

## startup

```{r}
library(tidyverse)
library(rstan)
library(ggrepel)
source("functions.R")
```

## get league info

## get games and league IDs

```{r}
games=readRDS("~/teaching/scoresway/rds/games.rds")
league_ids=readRDS("league_ids.rds")
league_ids
```

get country

```{r}
league_ids %>% arrange(the_country) %>% mutate(row=row_number()) %>% select(the_country, row, id) -> li
# View(li,"The leagues")
```

## enter number of desired country below

```{r league-no}
cno=31
prior_year=2020
li %>% slice(cno) -> v
v %>% unnest(id)
v %>% pull(id) %>% .[[1]] -> lnos
v %>% pull(the_country) -> country
v %>% pull(id) %>% .[[1]] -> id
iterations=10000
```


compare what we have to schedule

```{r}
lnos
games %>% filter(comp %in% lnos) %>% 
  arrange(desc(time_stamp)) %>% View("sched")
```





## make posterior

Change year if needed, then run whole chunk

### ```{r, echo=T, message=F, warning=F, results="hide"}
```{r} 
country_lower=tolower(country)
# prior_year=2019 # change only for Sweden/Nordic at end of year or for winter seasons during summer. Used several times below.
# prior_year=2019 # change only for Sweden/Nordic at end of year or for winter seasons during summer. Used several times below.
priorname=str_c(country_lower,prior_year,"_prior.csv")
prepostname=str_c(country_lower,"_pre_post.rds")
postname=str_c(country_lower,"_post.rds")
postname
prior=get_prior(priorname)
pre_post=readRDS(postname) # comment out these lines for a first time rating (ideally I want to use prior as pre_post)
saveRDS(pre_post, prepostname) # this one too
post=make_posterior_2(prior,lnos,iterations=iterations)
# post
# display(post,prior) %>% ggplot(aes(x=diff,y=sum, label=team))+geom_point()+geom_text_repel()+xlab("Openness")+ylab("Quality")
saveRDS(post,postname)
beepr::beep(2)
```





## display results

```{r}
prior %>% select(stan_id=id, sw_id) -> tab1
# tab1
tab2 <- name_lookup_table(games, lnos) # does this work for regular leagues?
# tab2 # I ignore the stan_id column because we are getting stan IDs from prior
tab1 %>% left_join(tab2, by="sw_id") %>% 
  select(stan_id=stan_id.x, name) -> name_table
# post
disp_old <- display_2(pre_post, name_table)
disp <- display_2(post, name_table)
disp %>% left_join(disp_old, by="name") %>% 
  mutate(new_rank=rank(desc(sum.x)),
         old_rank=rank(desc(sum.y)),
         change=old_rank-new_rank,
         delta=sum.x-sum.y,
         flag=case_when(
           abs(delta)>0.12 ~ "***",
           abs(delta)>0.08 ~ "**",
           abs(delta)>0.04 ~ "*",
           TRUE ~ ""
         )
         ) %>% 
        
  select(new_rank, old_rank, change, name, attack=o.x, defence=d.x, home=h.x, sum=sum.x, old_sum=sum.y, delta, flag) %>% 
  arrange(new_rank) -> display_update

games %>%  filter(comp %in% id) %>% filter(str_detect(score, " - ")) %>% 
  mutate(iso_week=iso_week_half(time_stamp)) %>% 
  filter(iso_week==max(iso_week)) %>% 
  select(t1, t2, score, time_stamp) %>% 
  separate(score, into=c("s1", "s2"), remove=F) %>% 
  mutate(score_rev=str_c(s2, s1, sep=" - ")) %>% 
  select(-s1, -s2) -> results
results %>% select(team=t1, opp=t2, score=score, time_stamp) %>% 
  mutate(venue="h") -> r1
results %>% select(team=t2, opp=t1, score=score_rev, time_stamp) %>% 
  mutate(venue="a") -> r2
bind_rows(r1, r2) -> ar
ar %>% left_join(tab2, by=c("team"="sw_id")) %>% 
  left_join(tab2, by=c("opp"="sw_id")) %>% 
  select(t1_name=name.x, t2_name=name.y, score, venue, time_stamp) -> all_results

display_update %>% 
  mutate_at(5:10, ~round(., 2)) %>% 
  left_join(all_results, by=c("name"="t1_name")) %>% 
  rename(opponent=t2_name) %>% View("new ratings")
```


```{r}
display_update
```

## graph (old)

```{r}
# namelc="ireland"
namelc=tolower(country)
display_2(post, name_table) %>% ggplot(aes(x=diff, y=sum, label=name))+geom_point()+geom_text_repel()+xlab("Openness")+ylab("Quality")+ggtitle(str_to_title(namelc))
```

```{r}
display_2(post, name_table)
```


## graph (2019-02-15) (needs fixing) name vs team probably


```{r}
graph_it(prior, pre_post, post) +  coord_fixed() + coord_flip()
```






## graph (old)

```{r}
# namelc="ireland"
namelc=tolower(country)
prior_year=2019
priorname=str_c(namelc, prior_year, "_prior.csv")
priorname
prior=get_prior(priorname)
postname=str_c(namelc, "_post.rds")
# prepostname=str_c(namelc, "_pre_post.rds")
post=readRDS(postname)
# pre_post=readRDS(prepostname)
display
display(post, prior) %>% ggplot(aes(x=diff,y=sum, label=team))+geom_point()+geom_text_repel()+xlab("Openness")+ylab("Quality")+ggtitle(str_to_title(namelc))
```

diagnose problems if any

```{r}
  games %>% filter(comp %in% lnos) -> current_games
  current_games %>% 
    filter(str_detect(score," - ")) %>% 
    separate(score,c("s1","s2"),sep=" - ",convert=T) %>% 
    select(t1_name,t2_name,s1,s2) -> gg
gg
```



## junk

## get leagues (probably not needed any more)

for given country and find division and season we want 
(type country name in.)
`nt` is date of *first* match of season.

run one line at a time

```{r}
country="Iceland"
(games %>% country_to_leagues(country) -> ll)
league_name=ll[6]
country_to_league_number(games,country,league_name)
lnos=47479
```

making a graph with pre-post and post

function to make df of posts

```{r}
post_df=function(prior,post_thing) {
  rat=extract(post_thing)
  tibble(id=1:nrow(prior),
         o=apply(rat$o,2,mean),
         d=apply(rat$d,2,mean),
         h=mean(rat$h)) %>% 
    left_join(prior,by=c("id")) %>% 
    select(team,id,o=o.x,d=d.x) %>% 
    mutate(sum=o+d,diff=o-d) %>% 
    arrange(desc(sum))
}
```

use this to make graph

```{r}
graph_it=function(prior, pre_post, post) {
  df1=post_df(prior, post)
  df0=post_df(prior, pre_post)
  bind_rows(df1, df0, .id="which") -> d # 1 is new, 2 is old
  d %>% group_by(which) %>% summarize(mm=mean(sum)) -> means
  d %>% left_join(means) %>% # look up right mean for "which"
    mutate(sum=sum-mm) %>% 
    mutate(which=ifelse(which==1, "now", "previous")) %>% 
    mutate(mylab=ifelse(which=="now", team, "")) -> d2
  ggplot(d2,aes(x=diff,y=sum,colour=which,label=mylab))+
    geom_point()+
    geom_path(aes(group=team),colour="blue", arrow=arrow(type="closed", ends="last", length=unit(0.1, "inches")))+
    geom_text_repel(colour="black")  
}
```

```{r}
graph_it(prior, pre_post, post)
```


playing

test it

```{r}
df1=post_df(prior, post)
df0=post_df(prior, pre_post)
bind_rows(df1, df0, .id="which") -> d # 1 is new, 2 is old
d
```

graph (based on Procrustes plot from D29)

```{r}
 ggplot(d,aes(x=diff,y=sum,colour=which,
                      label=team))+geom_point()+
    geom_line(aes(group=team),colour="green")+
    geom_text_repel()
```

need to standardize sums to have mean zero, and to not print name by previous. 

```{r}
d
(d %>% group_by(which) %>% summarize(mm=mean(sum)) -> means)
```

```{r}
(d %>% left_join(means) %>% mutate(sum=sum-mm) %>% 
   mutate(which=ifelse(which==1, "present", "past")) %>% 
   mutate(mylab=ifelse(which=="present", team, "")) -> d2)
```

graph take 2

```{r}
 ggplot(d2,aes(x=diff,y=sum,colour=which,
                      label=mylab))+geom_point()+
    geom_path(aes(group=team),colour="blue", arrow=arrow(type="closed", ends="first", length=unit(0.1, "inches")))+
    geom_text_repel(colour="black")
```



reference


```{r}
display=function(post,pre_post,prior) {
  rat=extract(post)
  tibble(id=1:nrow(prior),
         o=apply(rat$o,2,mean),
         d=apply(rat$d,2,mean),
         h=mean(rat$h)
  ) %>% 
    left_join(prior,by=c("id")) %>% 
    select(team,id,o=o.x,d=d.x) %>% 
    mutate(sum=o+d,diff=o-d) %>% 
    arrange(desc(sum))
}
```

### europe here

```{r}
country="europe"
iterations=1000
prior_year=2020
euro_leagues=tribble(
~league, ~y2020, ~y2019,
"euc", 54137, 48405,
"prsf", 54145, 48814,
"prf", 54144, 48968,
"2qr", 54138, 48406,
"3qr", 54139, 48407,
"po", 54143, 48411,
"gs", 54142, 48410, 
"8f", -1, 48408,
"qf", -1, 48412,
"sf", -1, 48413,
"f", -1, 48409,
"eue", 54012, 48815,
"1qr", 54005, 48744,
"2qr", 54006, 48745,
"3qr", 54007, 48746,
"po", 54011, 48750,
"gs", 54010, 48749,
"16f", -1, 48743,
"8f", -1, 48747,
"qf", -1, 48751,
"sf", -1, 48752,
"f", -1, 48748,
"albania", 54441, 48259,
"andorra", 53266, 47723,
"do", 53265, 47722, 
"armenia", 54074, 48391,
"do", 54897, -1,
"austria", 54164, 48232,
"do", 54161, 48721,
"azerbaijan", 53041, 47907,
"belarus", 51126, 45312,
"belgium", 53516, 48492,
"do", 53515, 48491,
"bosnia-herzegovina", 53769, 48238,
"bulgaria", 54087, 48506,
"do", 54083, 48502,
"croatia", 53842, 48337,
"cyprus", 53274, 47751,
"do 2", 53273, 47750,
"Czech rep", 54509, 48236,
"do", 54505, 48725,
"denmark", 54285, 48497,
"do", 54281, 48493,
"england", 53145, 48730,
"estonia", 50619, 44926,
"faroe is", 50678, 44303,
"finland", 51457, 44726,
"france", 53638, 48044,
"georgia", 51162, 45330, 
"germany", 53499, 47657,
"gibraltar", 53435, 48922,
"do", 55353, -1, 
"greece", 52837, 47479,
"do 2", 53558, -1, 
"hungary", 53519, 48855,
"iceland", 50336, 44141,
"israel", 53556, 48018,
"italy", 54890, 48235,
"kazakhstan", 50618, 44925,
"kosovo", 53364, 48248,
"latvia", 50609, 44854,
"liechtenstein", 52832, 47474,
"ditto 2", 52833, 47475,
"ditto 3", 52834, 47477,
"ditto 4", 52835, 47478,
"ditto 5", 52836, 47476,
"lithuania", 50746, 45171,
"do 2", 50745, 45170,
"luxembourg", 53290, 48021,
"malta", 52902, 47326,
"moldova", 50991, 45337,
"montenegro", 53867, 48414,
"netherlands", 54058, 47971,
"north macedonia", 53864, 48390,
"northern ireland", 53081, 47720,
"do", 53078, 47717,
"norway", 50939, 45310, 
"poland", 53505, 48402,
"do 2", 53504, 48401,
"portugal", 53517, 47741,
"rep ireland", 50855, 45768,
"romania", 54245, 48876,
"do 2", 54244, 48875,
"russia", 53628, 47835,
"san marino", 54124, 49261,
"do 2", -1, 49257,
"do 4", -1, 49253,
"do 5", -1, 49254,
"do 6", -1, 49252,
"scotland", 53513, 47730,
"do 2", 53514, 47731,
"serbia", 53321, 48023,
"do", 53320, 48022,
"slovakia", 53771, 48246,
"do", 53770, 48243,
"slovenia", 53678, 48127,
"spain", 53502, 47983,
"sweden", 50679, 44851,
"switzerland", 54327, 48769,
"turkey", 53866, 48404,
"ukraine", 54067, 47821, 
"ditto 2", 54066, 47820,
"wales", 53473, 48196,
"ditto 2", 53474, 48197
)
```

```{r}
euro_leagues %>% filter(y2020>0) %>% 
  pull(y2020) -> lnos
lnos
```


### facup and leagues (do I need to fill out levels 6, 7 and 8?)

```{r}
country="facup"
iterations=1000
prior_year=2020
facup_leagues=tribble(
  ~league, ~y2020, ~y2019, ~y2018,
  "p", 53384, 48359, 42260,
  "pr", 53385, 48360, 42261,
  "q1", 53366, 48339, 42240,
  "q1r", 53367, 48340, 42241,
  "q2", 53370, 48343, 42244,
  "q2r", 53371, 48344, 42245,
  "q3", 53374, 48347, 42248,
  "q3r", 53375, 48348, 42249,
  "q4", 53378, 48351, 42252,
  "q4r", 53379, 48352, 42253,
  "r1", 53365, 48338, 42239,
  "r1r", 53368, 48341, 42242,
  "r2", 53369, 48342, 42243,
  "r2r", 53372, 48345, 42246,
  "r3", 53373, 48346, 42247,
  "r3r", 53376, 48349, 42250,
  "r4", 53377, 48350, 42251,
  "r4r", 53380, 48353, 42254,
  "r5", 53381, 48354, 42255,
  "r5r", -1, -1, 42256,
  "r6", -1, 48361, 42262,
  "sf", -1, 48362, 42263,
  "f", -1, 48358, 42259,
  "l1", 53145, 48730, 41547,
  "l2", 53782, 48274, 42070,
  "l3", 53677, 48115, 41820,
  "l4", 53874, 48450, 42073,
  "l5", 53131, 48734, 41651,
  "l6a", 53161, 47691, 41337,
  "l6b", 53165, 47695, 41340,
  "l7a", 53153, 47526, 40941, 
  "l7b", 53150, 47523, 40938,
)
```








grab the ones to make posterior out of

```{r}
facup_leagues %>% filter(y2020>0) %>% pull(y2020) -> lnos
```