---
title: "rugby league"
output: html_notebook
---

## packages

```{r}
library(tidyverse)
library(rstan)
library(ggrepel)
source("functions.R")
```


## priors

use 2019 data (update "games")

fix up 2019 league 1

```{r}
games=readRDS("~/Documents/flashscore/games.rds")
# View(games)
```

```{r}
games %>% filter(comp==201901) -> games
# View(games)
```


```{r}
lnos=201901
(games %>% comp_games(lnos) -> gg)
(gg %>% team_lookup_table() -> lookup_table)
```

```{r}
make_stan(games,lookup_table)
```


```{r}
games %>% 
  make_stan_prior(lnos,"rl201903.csv") 
beepr::beep(3)
```

go back and read in the original games

```{r}
games=readRDS("~/Documents/flashscore/games.rds")
(games %>% comp_games(202003)
  %>% team_lookup_table() -> lt1)
(games %>% comp_games(201903) %>% team_lookup_table() -> lt2)
lt2 %>% anti_join(lt1,by="team") # removed from
lt1 %>% anti_join(lt2,by="team") # added to

```

make this year's prior

now edit last year's ratings to be this year's prior, removing the promoted and relegated teams, and replacing them with the ones that went down and up last year.

having replaced up and down names with u and d:

```{r}
prior_name="rl202003_prior.csv"
pr=read_csv(prior_name)
pr %>% filter(str_length(team)==1) %>% 
  group_by(team) %>% 
  summarize(m=mean(o+d)/2)
```

## posterior

read games

```{r}
games=readRDS("~/Documents/flashscore/games.rds")
games
```


### get names of everything (change league number)

```{r}
lnos=202003
country="rl"
country_lower="rl"
prior_year=lnos # change only for Sweden/Nordic at end of year
priorname=str_c(country_lower,lnos,"_prior.csv")
prepostname=str_c(country_lower,lnos,"_pre_post.rds")
postname=str_c(country_lower,lnos,"_post.rds")
prior=get_prior(priorname)
pre_post=readRDS(postname) # comment out these lines for a first time rating (ideally I want to use prior as pre_post)
saveRDS(pre_post, prepostname) # this one too
```


make posterior

```{r}
post=make_posterior(prior,lnos)
# display(post,prior) %>% ggplot(aes(x=diff,y=sum, label=team))+geom_point()+geom_text_repel()+xlab("Openness")+ylab("Quality")
saveRDS(post,postname)
```

```{r}
post
```


```{r}
display(post, prior) %>% View("posterior")
```

changes (needs, um, changing. Think about priors and if I have anything like sw_id and if it matters.)

```{r}
df1=post_df(prior,post)
df0=post_df(prior,pre_post)
df1 %>% left_join(df0, by=c("team"="team")) %>% 
  mutate(sum.x=sum.x-mean(sum.x),
         sum.y=sum.y-mean(sum.y)) %>% 
  mutate(delta_sum=sum.x-sum.y,
         delta_diff=diff.x-diff.y) %>% 
  mutate(pr_sum=percent_rank(sum.x),
         pr_diff=percent_rank(diff.x)) %>% 
  select(team, pr_sum, pr_diff, delta_sum, delta_diff) -> z
z %>% arrange(desc(abs(delta_diff)))
z %>% arrange(desc(abs(delta_sum)))
```

### graph (old)

```{r, fig.width=8, fig.height=8}
display(post, prior) %>% ggplot(aes(x=diff,y=sum, label=team))+geom_point()+geom_text_repel()+xlab("Openness")+ylab("Quality")+ggtitle(str_to_title(lnos))
```


take a look at post (for just looking, use pre-post)

```{r}
post_df(prior, pre_post)
```

## previous predictions alongside actual game results


```{r}
p_names=list.files("~/Documents/ratings","preds.*.rds")
enframe(p_names) %>% arrange(desc(value))
```

```{r}
which=32
pred_file=p_names[which]
preds=readRDS(pred_file)
preds %>% mutate(key=str_c(date, home_team, away_team, sep=" ")) -> preds
```

```{r}
games_2=readRDS("~/Documents/flashscore/games_2.rds")
games_2
games_2 %>% mutate(key=str_c(date, t1_name, t2_name, sep=" ")) -> games_2
preds %>% left_join(games_2, by="key") %>% select(date=date.x, name, home_team, prediction, score, away_team) %>% drop_na() %>% View("predictions-actual")
```

## post?

```{r}
prior
post=pre_post
m=extract(post)
m$o
m$d
m$h
r=with(m, data.frame(o=o,d=d,h=h))
r
```

```{r}
t1=7
t2=10
no1=str_c("o.", t1)
no2=str_c("o.", t2)
nd1=str_c("d.", t1)
nd2=str_c("d.", t2)
no1
nd2
```

```{r}
n=nrow(r)
r %>% select(o1=no1, o2=no2, d1=nd1, d2=nd2, h) %>% 
  mutate(l1=exp(o1-d2+h), l2=exp(o2-d1)) %>% 
  mutate(s1=rpois(n,l1), s2=rpois(n,l2)) %>% 
  select(s1, s2) -> sim
sim %>% count(s1, s2) -> counts
ggplot(counts, aes(x=2*s1, y=2*s2, fill=n)) + geom_tile() + geom_abline(slope=1, intercept=0)
sim %>% 
  mutate(result=case_when(s1>s2 ~ "W",
                          s1<s2 ~ "L", 
                          TRUE  ~ "D")) %>% 
  count(result) %>% 
  mutate(proportion=n/sum(n))
counts %>% arrange(desc(n)) %>% mutate(s1=2*s1, s2=2*s2)
sim %>% mutate_all(~2*.) %>% 
  summarize_all(~median(.))
```

```{r}
sim %>% gather(team, score, everything()) %>% 
  ggplot(aes(x=2*score))+geom_bar(bins=20)+facet_grid(team~.)
```


make a function out of some of this

## fixtures and predictions

so I have fixtures:

```{r}
fixtures=readRDS("~/Documents/flashscore/fixtures.rds")
fixtures 
```

this needs fixing up (the function), possibly because post_df needs work (or the output from it does)

```{r}
pred_league=function(lnos, fixtures) {
  print(lnos)
  country_lower="rl"
  priorname=str_c(country_lower,lnos,"_prior.csv")
  postname=str_c(country_lower,lnos,"_post.rds")
  prior=get_prior(priorname)
  post=readRDS(postname)
  post_df(prior,post) %>% select(team, o, d) -> r
  return(post_df(prior, post))
  fixtures %>% filter(comp==lnos) %>% select(date, t1_name, t2_name) -> f
  f %>% left_join(r, by=c("t1_name"="name")) %>% 
  left_join(r, by=c("t2_name"="name")) %>% 
  mutate(hs=2*round(exp(o.x-d.y),0),
         as=2*round(exp(o.y-d.x),0),
         prediction=str_c(hs, " - ", as)) %>% 
  select(date, home_team=t1_name, prediction, away_team=t2_name) 
}
```

testing

```{r}
pred_league(201901, fixtures)
```


or,  grab posterior distribution and simulate (maybe make heat map)

make all the leagues

```{r}
leagues <- tribble(
  ~name, ~lno,
  "Super League", 201901,
  "Championship", 201902,
  "League 1", 201903,
  "NRL", 201911
)
```

```{r}
leagues %>% mutate(preds=map(lno, ~pred_league(., fixtures))) %>% 
  unnest() %>% arrange(date) -> preds
today=Sys.Date()
saveRDS(preds, str_c("preds.",today,".rds"))
```

old predictions

```{r}
list.files("~/Documents/ratings","preds.*.rds")
```

```{r}
fname="preds.2019-09-01.rds"
preds=readRDS(fname)
preds %>% mutate(key=str_c(home_team, away_team, date)) -> preds
preds 
```

## with actual?

```{r}
games=readRDS("~/Documents/flashscore/games_2.rds")
games %>% mutate(key=str_c(t1_name, t2_name, date)) -> games
preds %>% 
  left_join(games, by="key") %>% 
  select(date=date.x,home_team, away_team, prediction, score) %>% 
  View("rugby league prediction-actual")
```

### debugging

```{r}
league_number=201902
  games %>% filter(comp %in% league_number) -> current_games
  current_games %>% 
    filter(str_detect(score," - ")) %>% 
    separate(score,c("s1","s2"),sep=" - ",convert=T) %>% 
    select(t1_name,t2_name,s1,s2) -> gg
  if (nrow(gg)==0) { # no games yet
    # pull off first row from current_games and fill in 2-1 score for it
    current_games %>% slice(1) %>% 
      mutate(s1=2, s2=1) %>% 
      select(t1_name,t2_name,s1,s2) ->
      gg
  }
  # the below should squeal if I got any team names wrong in prior
  gg %>% left_join(prior,by=c("t1_name"="team")) %>% 
    left_join(prior,by=c("t2_name"="team")) %>%
    select(t1=id.x,t2=id.y,s1,s2) ->
    post_stan
  gg
post_stan
```

