---
title: "sweden"
output: html_notebook
---

attempting to do last season's Sweden ratings; not saving anything


## packages

```{r}
library(tidyverse)
library(rstan)
```

## get games

```{bash}
ls -l ~/teaching/scoresway/rds/games.rds
```
that's it:

```{r}
games=readRDS("~/teaching/scoresway/rds/games.rds")
games
```

get comp and comp_name. These are the first lines that match.

```{r}
games %>% select(comp, comp_name, time_stamp, country) %>% 
  group_by(comp) %>% 
  summarize(nm=min(comp_name),nt=min(time_stamp),nc=min(country)) %>% 
  filter(str_detect(nc, "Sweden"), str_detect(nm, "Allsvenskan"))
```

a function to return all the different leagues matching input country

```{r}
country_to_leagues=function(games,country) {
  games %>% select(comp, comp_name, time_stamp, country) %>% 
    group_by(comp) %>% 
    summarize(nm=min(comp_name),nt=min(time_stamp),nc=min(country)) %>% 
    filter(str_detect(nc, country)) %>% pull(nm) %>% unique()
}
```

```{r}
games %>% country_to_leagues("Sweden")
```

a function to return all the seasons matching country and league name

```{r}
country_to_league_number=function(games,country,league_name) {
  games %>% select(comp, comp_name, time_stamp, country) %>% 
    group_by(comp) %>% 
    summarize(nm=min(comp_name),nt=min(time_stamp),nc=min(country)) %>% 
    filter(str_detect(nc, country),str_detect(nm, league_name))
}
```

```{r}
country_to_league_number(games,"Sweden","Allsvenskan")
```

try another one

```{r}
country_to_leagues(games,"France")
country_to_league_number(games,"France","Ligue 1")
```

```{r}
comp_games=function(games,comp_number) {
  games %>% filter(comp==comp_number)
}
```

```{r}
team_lookup_table=function(the_games) { # the_games is output from comp_games 
  team_names=with(the_games,c(t1_name,t2_name))
  team_names=sort(unique(team_names))
  tibble(team=team_names) %>% mutate(id=row_number())
}
```

testing

```{r}
games %>% comp_games(39293) %>% team_lookup_table() # I probably shouldn't do it like this though
```

better:

```{r}
(games %>% comp_games(39293) -> gg)
(gg %>% team_lookup_table() -> lookup_table)
```


get nt=number of teams (as #rows in this) sometime

Look up team names by number

```{r}
make_stan=function(the_games,lookup_table) {
  the_games %>% 
    left_join(lookup_table,by=c("t1_name"="team")) %>% 
    left_join(lookup_table,by=c("t2_name"="team")) %>% 
    separate(score,c("s1","s2"),convert=T) %>% 
    select(t1=id.x, t2=id.y, s1, s2)
}
make_stan(gg,lookup_table)
```


```{r}
make_stan_prior_data=function(games,comp_id) {
  games %>% comp_games(comp_id) -> gg
  gg %>% team_lookup_table() -> lookup_table
  gg_stan=make_stan(gg, lookup_table)
  with(gg_stan,list(
    nt=nrow(lookup_table),
    ng=nrow(gg_stan),
    x=cbind(t1,t2),
    y=cbind(s1,s2),
    prior_o_mean=rep(0,nt),
    prior_o_sd=rep(1,nt),
    prior_d_mean=rep(0,nt),
    prior_d_sd=rep(1,nt),
    prior_h_mean=0,
    prior_h_sd=1
  ))
}
games %>% 
  make_stan_prior_data(39293) ->
spd
str(spd)
```

get the compiled model

```{r}
(p.sc=readRDS("psc.rds"))
```



now run the model

```{r}
sweden2017=sampling(p.sc,sweden_data)
sweden2017
```

extract stuff and get means

```{r}
rat=extract(sweden2017)
str(rat)
tibble(id=1:nt,
       o=apply(rat$o,2,mean),
       os=apply(rat$o,2,sd),
       d=apply(rat$d,2,mean),
       ds=apply(rat$d,2,sd),
       h=mean(rat$h),
       hs=sd(rat$h)
       ) %>% 
  left_join(teams) %>% 
  select(team_names,id,everything()) %>% 
  write_csv("sweden2017.csv")
```

make csv of mean and SD of ratings for each team this year (above)
edit to include promoted teams



find teams not in it this season

this season is 44851

```{r}
games %>% 
  filter(comp==44851) ->
d
teams_now=with(d,c(t1_name,t2_name))
teams_now=sort(unique(teams_now))
dd=tibble(team_names=teams_now)
dd %>% anti_join(teams) # teams not in it last year
teams %>% anti_join(dd) # teams in it last year but not this.
```


make this into a function
 
perhaps start with competition id

## league table

get games

```{r}
games %>% filter(comp==44851) %>% 
  filter(str_detect(score," - "))%>% 
  separate(score,c("s1","s2"),sep=" - ",convert=T) %>%
  select(t1_name,t2_name,s1,s2) %>% 
  mutate(result=case_when(
    s1>s2 ~ "W",
    s1==s2 ~ "D",
    s1<s2 ~ "L",
    TRUE ~ "X"
  )) %>% 
  mutate(pt1=case_when(
    result=="W" ~ 3,
    result=="D" ~ 1,
    result=="L" ~ 0
  )) %>% 
  mutate(pt2=case_when(
    result=="W" ~ 0,
    result=="D" ~ 1,
    result=="L" ~ 3
  )) -> 
d
d
```

info for home teams

```{r}
d %>% group_by(t1_name) %>% 
  summarize(P=n(),
            GD=sum(s1)-sum(s2),
            Pt=sum(pt1)
            ) -> 
tab1
d %>% group_by(t2_name) %>% 
  summarize(P=n(),
            GD=sum(s2)-sum(s1),
            Pt=sum(pt2)
            ) -> 
tab2
tab1 %>% left_join(tab2,by=c("t1_name"="t2_name")) %>% 
  mutate(P=P.x+P.y,GD=GD.x+GD.y,Pt=Pt.x+Pt.y) %>% 
  select(t1_name,P,GD,Pt) %>% 
  arrange(desc(Pt),desc(GD))
```

function to make league table from comp number

```{r}
table_of=function(games,lg_number) {
  games %>% 
    filter(comp==lg_number) %>% 
  filter(str_detect(score," - "))%>% 
  separate(score,c("s1","s2"),sep=" - ",convert=T) %>%
  select(t1_name,t2_name,s1,s2) %>% 
  mutate(result=case_when(
    s1>s2 ~ "W",
    s1==s2 ~ "D",
    s1<s2 ~ "L",
    TRUE ~ "X"
  )) %>% 
  mutate(pt1=case_when(
    result=="W" ~ 3,
    result=="D" ~ 1,
    result=="L" ~ 0
  )) %>% 
  mutate(pt2=case_when(
    result=="W" ~ 0,
    result=="D" ~ 1,
    result=="L" ~ 3
  )) -> 
d
d %>% group_by(t1_name) %>% 
  summarize(P=n(),
            GD=sum(s1)-sum(s2),
            Pt=sum(pt1)
            ) -> 
tab1
d %>% group_by(t2_name) %>% 
  summarize(P=n(),
            GD=sum(s2)-sum(s1),
            Pt=sum(pt2)
            ) -> 
tab2
tab1 %>% left_join(tab2,by=c("t1_name"="t2_name")) %>% 
  mutate(P=P.x+P.y,GD=GD.x+GD.y,Pt=Pt.x+Pt.y) %>% 
  select(t1_name,P,GD,Pt) %>% 
  arrange(desc(Pt),desc(GD))
}
```

test it

```{r}
table_of(games,39292)
```

hand-edited last year's ratings to make priors for this year. Now use as priors, doubling the SDs

```{r}
sweden_prior=read_csv("sweden2018_prior.csv")
sweden_prior
```

## construct data list

results with team numbers

```{r}
games %>% filter(comp==44851) %>% 
  filter(str_detect(score," - "))%>% 
  separate(score,c("s1","s2"),sep=" - ",convert=T) %>% 
  select(t1_name,t2_name,s1,s2) %>% 
  left_join(sweden_prior,by=c("t1_name"="team_names")) %>% 
  left_join(sweden_prior,by=c("t2_name"="team_names")) %>% 
  select(t1=id.x,t2=id.y,s1,s2) ->
sweden_stan
```

now make list

```{r}
sweden_data=with(sweden_stan,list(
  nt=nrow(sweden_prior),
  ng=nrow(sweden_stan),
  x=cbind(t1,t2),
  y=cbind(s1,s2),
  prior_o_mean=sweden_prior$o,
  prior_o_sd=2*sweden_prior$os,
  prior_d_mean=sweden_prior$d,
  prior_d_sd=2*sweden_prior$ds,
  prior_h_mean=sweden_prior$h[1],
  prior_h_sd=sweden_prior$hs[1]
))
str(sweden_data)
p.sc=readRDS("psc.rds")
p.sc
```

```{r}
sweden_now=sampling(p.sc,data=sweden_data)
```

```{r}
sweden_now
```

```{r}
rat=extract(sweden_now)
str(rat)
tibble(id=1:nrow(sweden_prior),
       o=apply(rat$o,2,mean),
       d=apply(rat$d,2,mean),
       h=mean(rat$h)
       ) %>% 
  left_join(sweden_prior,by=c("id")) %>% 
  select(team_names,id,o=o.x,d=d.x) %>% 
  mutate(sum=o+d,diff=o-d) %>% 
  arrange(desc(sum))
```

